---
title: "Generating Errors"
author: "Tito"
date: "2022-11-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Generate Regular Point Count Data

In the other .RMD notebook "Process to Create Point Count Data" regular point count
data has been made. The next step is to now create a latent matrix and use that to
create the errors. An additional variable alpha has been created to represent
the probability of a bird moving.


## Set Variables
```{r}
source("./Functions/generatePointCount.R")
```


```{r}
p_1m = 0.6 # needs to be the same p_1m I use to generate summarized count data
maxMinute = 5
psi=0.6
n_locations = 50
lambda = 15 # get this number from some data?
alpha = 0.3
```


## Generate Population Data
```{r}
set.seed(1)
locationID = 1:n_locations
#N_i = rpois(n_locations,lambda)
N_i = rnbinom(n_locations, mu = lambda, size= 1)
```


```{r}
# view results
result = returnData(N_i,p_1m,maxMinute, seed = 1)
dfRegular_Cumulative = result[[1]]
dfRegular = result[[2]]
dfRegular_Summarized = result[[3]]


head(dfRegular_Cumulative)
head(dfRegular)
head(dfRegular_Summarized)
```

```{r}
generateLatentMatrix = function(Ni,alpha,maxMinute, seed = NULL){
  if(!is.null(seed))
    set.seed(seed)
  
  if(Ni == 0){
    result = NULL
  } else {
    individuals = 1:Ni
    latent_1 = as.data.frame(individuals)
    
    for(time in 1:(maxMinute-1)){
      
      # the time variable in this case will represent what happens *after* that time number
      # so data in column "1" represents if the bird moved *after* time 1
      
      # I'm saying each individual is moves or not (bernoulli) w.p. alpha
      sample_movement = as.data.frame(rbinom(nrow(latent_1),1,alpha))
      colnames(sample_movement)=time
      latent_1 = cbind(latent_1,sample_movement)
      
    }
    result = latent_1
  }
  
  return(result)
  
}
```


```{r}
latentMatrix = generateLatentMatrix(4,0.5,5,1)
latentMatrix
```

Use this as a case to check! This shows me that I am splitting correctly. 
However, I need to implement logic to drop the extra row that is split upon 
if more than one movement happens (ie in this example row 1_3b would need to be dropped)
```{r}
# row in 1:nrow(latentMatrix
finalDF = matrix(NA,0,0)
for(row in 1:1){

  # select a row to isolate detection history and movement history for an individual
  originalHistory = as.numeric(dfRegular[row,2:(maxMinute+1)]) # this will be updated as we loop through time intervals and make splits
  latentVector = as.numeric(latentMatrix[row,2:(maxMinute)])
  
  # figure out which columns have movement
  timeMovement = which(latentVector==1)
  
  # loop through the time intervals with movement and perform splits at each 
  # time interval
  for(time in timeMovement){
    # keep old vector values from before movement
    new_time = numeric(5)
    new_time[1:time] = originalHistory[1:time]
    
    # add new individual and give them remaining counts
    new_time_move = numeric(5)
    new_time_move[-c(1:time)] = originalHistory[-c(1:time)]
    
    # bind the two vectors together
    tempDF = rbind(new_time, new_time_move)
    colnames(tempDF) = 1:5
    rownames(tempDF) = NULL
    tempDF = as.data.frame(tempDF)
    tempDF$identifier = paste0(row,"_",time,letters[1:2])
    
    # update originalHistory vector so that it is the vector of the "new" individual
    originalHistory = new_time_move

    # return two vectors!
    if(nrow(finalDF)==0){
      finalDF = tempDF
    } else {
      finalDF = bind_rows(finalDF, tempDF)
    }

   }
   print(finalDF)
}
```

Added a counter so that the last split is not removed but other splits will be!
```{r}
# row in 1:nrow(latentMatrix
finalDF = matrix(NA,0,0)
for(row in 1:1){

  # select a row to isolate detection history and movement history for an individual
  originalHistory = as.numeric(dfRegular[row,2:(maxMinute+1)]) # this will be updated as we loop through time intervals and make splits
  latentVector = as.numeric(latentMatrix[row,2:(maxMinute)])
  
  # figure out which columns have movement
  timeMovement = which(latentVector==1)
  counter = 0 # to count how many iterations so far
  
  # loop through the time intervals with movement and perform splits at each 
  # time interval
  for(time in timeMovement){
    # keep old vector values from before movement
    new_time = numeric(5)
    new_time[1:time] = originalHistory[1:time]
    
    # add new individual and give them remaining counts
    new_time_move = numeric(5)
    new_time_move[-c(1:time)] = originalHistory[-c(1:time)]
    
    # bind the two vectors together
    tempDF = rbind(new_time, new_time_move)
    colnames(tempDF) = 1:5
    rownames(tempDF) = NULL
    tempDF = as.data.frame(tempDF)
    tempDF$identifier = paste0(row,"_",time,letters[1:2])
    
    # update originalHistory vector so that it is the vector of the "new" individual
    originalHistory = new_time_move
    counter = counter + 1 # update counter 
    
    # add case to remove redundant rows
    if(counter<length(timeMovement)){
      tempDF = tempDF %>%
        filter(identifier != paste0(row,"_",time,"b"))
    }

    # return two vectors!
    if(nrow(finalDF)==0){
      finalDF = tempDF
    } else {
      finalDF = bind_rows(finalDF, tempDF)
    }
    

   }
   print(finalDF)
}
```

Generalize into a function, and take care of cases where no movement occurs.
```{r}
generateErrors = function(dfPointCount,latentMatrix,maxMinute){
  
  # handle the cases where latentMatrix is NULL (because population is zero at that location)
  if(is.null(latentMatrix)){
    finalDF = NULL
  } else {
    finalDF = matrix(NA,0,0)
    for(row in 1:nrow(latentMatrix)){

    # select a row to isolate detection history and movement history for an individual
    locationID = dfPointCount[row,]$locationID
    originalHistory = as.numeric(dfPointCount[row,2:(maxMinute+1)]) # this will be updated as we loop through time intervals and make splits
    latentVector = as.numeric(latentMatrix[row,2:(maxMinute)])
    
    # figure out which columns have movement
    timeMovement = which(latentVector==1)
    counter = 0 # to count how many splits have happened so far
    
    # handle case where no movement occurs
    if(length(timeMovement)==0){
      tempDF = as.data.frame(originalHistory)
      tempDF = t(tempDF)
      rownames(tempDF) = NULL
      colnames(tempDF) = 1:maxMinute
      tempDF = as.data.frame(tempDF)
      tempDF$individuals = row
      tempDF$individuals = as.character(tempDF$individuals)
      tempDF$locationID = locationID
      
      # return two vectors!
      if(nrow(finalDF)==0){
        finalDF = tempDF
      } else {
        finalDF = bind_rows(finalDF, tempDF)
      }
      
    } else {
      # loop through the time intervals with movement and perform splits at each 
      # time interval
      for(time in timeMovement){
        
        # keep old vector values from before movement
        new_time = numeric(maxMinute)
        new_time[1:time] = originalHistory[1:time]
        
        # add new individual and give them remaining counts
        new_time_move = numeric(maxMinute)
        new_time_move[-c(1:time)] = originalHistory[-c(1:time)]
        
        # bind the two vectors together
        tempDF = rbind(new_time, new_time_move)
        colnames(tempDF) = 1:maxMinute
        rownames(tempDF) = NULL
        tempDF = as.data.frame(tempDF)
        tempDF$individuals = paste0(row,"_",time,letters[1:2])
        tempDF$locationID = locationID
        
        # update originalHistory vector so that it is the vector of the "new" individual
        originalHistory = new_time_move
        counter = counter + 1 # update counter 
        
        # add case to remove redundant rows
        if(counter<length(timeMovement)){
          tempDF = tempDF %>%
            filter(individuals != paste0(row,"_",time,"b"))
        }
    
        # return two vectors!
        if(nrow(finalDF)==0){
          finalDF = tempDF
        } else {
          finalDF = bind_rows(finalDF, tempDF)
        }
      } 
    }
  }
}
  
  # remove duplicated rows 
  return(finalDF)
}
```


```{r}
generateErrors(dfRegular[1:1,], latentMatrix[1:1,], maxMinute)
```

```{r}
head(dfRegular)
generateLatentMatrix(0,0.3,5,1)
```

