---
title: "Process to Create Point Count Data"
author: "Tito"
date: "2022-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(data.table)
library(matrixStats)
source("./Functions/generatePointCount.R")
```


# Process to create Point Count Data

## Set Variables
```{r}
p_1m = 0.1
maxMinute = 10
colsUse = paste0("Minute",1:maxMinute)
n_locations = 50
lambda = 15 # get this number from some data?
```


## Generate Population Data
```{r}
set.seed(1)
locationID = 1:n_locations
#N_i = rpois(n_locations,lambda)
N_i = rnbinom(n_locations, mu = lambda, size= 1)
dfPop = as.data.frame(cbind(locationID, N_i))
head(dfPop)
```

Point count data in long format.
```{r}
dfCount = generateDetects(dfPop$N_i, p_1m,maxMinute,10)
head(dfCount)
```

Point count data in wide format, with zero entries dropped
```{r}
dfWide = detectsToCapHist(dfCount)

# add total individuals at this location 
dfWide = left_join(dfWide, dfPop, by = "locationID")
head(dfWide)
```



## Results for Bin(Ni,p)

Add summarized data and view cumulative data
```{r}
dfPop = left_join(dfPop, formatData(dfCount, isSummarised = TRUE), by = "locationID")
dfPop$C_i = replace_na(dfPop$C_i,0) # fill NA that occurs when N_i is 0
head(dfPop)
```

```{r}
dfCumulative = formatData(dfCount, isCumulative = TRUE)
dfCumulative = left_join(dfCumulative, dfPop[,c("locationID","N_i")], by = "locationID")
head(dfCumulative)
```


# Graphing

Some graphs to see if generated results make sense

Distribution of counts is similar to the population - but range is not as large 
which is how it should be. Looks reasonable. 
```{r}
dfPop %>%
  pivot_longer(cols = c("N_i","C_i"), names_to = "CountType", values_to = "Value")%>%
  ggplot()+ 
  geom_density(aes(x=Value, color = CountType, fill = CountType),alpha=0.7)+ 
  facet_wrap(~CountType,scales="free")+ 
  theme_minimal()+ 
  labs(title = "Distribution of Site Specific Population Values Against Point Counts",
       subtitle = "Point counts follow a binomial distribution")
```

Similar to the probability of detecting an individual at least once
```{r}
mean(dfPop$C_i/dfPop$N_i,na.rm=TRUE)
PredictProbability(maxMinute,p_1m)
```

Graph shows the distribution of the number of times each individual has been counted
by the end of m minutes. Graph makes sense seeing as more time passes, number
of individuals that have been counted 1 or 2 times grows --> looks like a multinomial 
distribution. 

```{r}
dfCumulative%>%
  pivot_longer(cols = c(colsUse), names_to = "Time", values_to = "Count")%>%
ggplot()+ 
  geom_density(aes(x=Count, fill = Time, color = Time),alpha=0.6)+ 
  facet_wrap(~Time,scales= "free")+ 
  theme_minimal()+
  labs(title= "Cumulative Point Counts by Time Interval")
```

## Checking Detection Probability 

If c/p is an estimator for $\hat{p}$ then I want to see how it changes as 
the time interval changes. 
```{r}
CountByTimeInterval=dfCumulative %>%
  pivot_longer(cols = c(colsUse), names_to = "TimeInterval", values_to = "PointCount")%>%
  group_by(locationID, individual, TimeInterval)%>%
  summarise(TotalCount = sum(PointCount))

CountByTimeInterval$TotalCount= ifelse(CountByTimeInterval$TotalCount>=1,1,0)
CountByTimeInterval=CountByTimeInterval %>%
  group_by(locationID, TimeInterval)%>%
  summarise(FinalCount = sum(TotalCount))
# add total location population
CountByTimeInterval = left_join(CountByTimeInterval, dfPop, by = "locationID")
head(CountByTimeInterval)
```

Find the implied p_hat of the data. 
```{r}
CountByTimeInterval = CountByTimeInterval %>%
  mutate(p_hat = FinalCount/N_i)
CountByTimeInterval$TimeInterval = as.numeric(gsub('Minute', '', CountByTimeInterval$TimeInterval))
head(CountByTimeInterval)
```

```{r}
# find theoretical probability of being detected at least once
summaryPhat = CountByTimeInterval %>%
  group_by(TimeInterval)%>%
  summarise(p_data = mean(p_hat))
summaryPhat$TimeInterval = as.numeric(summaryPhat$TimeInterval)
summaryPhat$p_theoretical = PredictProbability(summaryPhat$TimeInterval, p_1m)
summaryPhat
```


```{r}
summaryPhat %>%
  pivot_longer(cols=c("p_data","p_theoretical"),names_to = "Type",values_to = "p" )%>%
ggplot()+
  geom_point(aes(x=TimeInterval, y = p,color=Type))+ 
  geom_line(aes(x=TimeInterval, y = p,color= Type))+
  theme_minimal()+ 
  labs(title = "Comparison of Estimated Detection Probability and Theoretical Probability",
       subtitle = "Theoretical probability is the probability of being detected at least once.")
```

