---
title: "Process to Create Errors in Point Count Data"
author: "Tito"
date: "2022-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set Variables
```{r}
library(tidyverse)
```


```{r}
source("./Functions/generatePointCount.R")
source("./Functions/generatePointCountErrors.R")
```


#Q:

How to deal with the case where the bird is only detected once - allow double counting?
```{r}
p_1m = 0.6 # needs to be the same p_1m I use to generate summarized count data
maxMinute = 5
psi=0.6
n_locations = 50
lambda = 15 # get this number from some data?
alpha = 0.3
```


## Generate Population Data
```{r}
set.seed(1)
locationID = 1:n_locations
#N_i = rpois(n_locations,lambda)
N_i = rnbinom(n_locations, mu = lambda, size= 1)
```


```{r}
# view results
result = returnData(N_i,p_1m,maxMinute, seed = 1)
dfRegular_Cumulative = result[[1]]
dfRegular = result[[2]]
dfRegular_Summarized = result[[3]]
dfRegular_Summarized$locationID = 1:n_locations


head(dfRegular_Cumulative)
head(dfRegular)
head(dfRegular_Summarized)
```

```{r}
# for reference so we can see what latent matrix is used to create errors
latentMatrixUsed = generateLatentMatrix(dfRegular,alpha,maxMinute,1)
latentMatrixUsed
```

I have numbered the individuals so that it is easy to identify where the error came
from and what the new detection histories mean. 

- If the individual is represented by a single integer, it means there was no 
movement at all in the latent matrix
- If the individual is denoted following the form (1_2a or 1_2b - so the form number_number_letter) it means that 
individual 1 moved immediately after time 2. The a represents the detection history before 
the split, and the b represents the new individual made after the double counting occurred.
*I assume each movement = double counting error*
- If there is an individual with the format 2_3 it means the second individual was
split at time 3, but because the split would've produced a new (or old) history with 
all 0's in the detection history, we ignore that split because it would be a labelling
error instead of a double counting error. 
  - For example if we look at individual 3, we can see that movement happened after 
  time 2 and time 4.
  - the 3_2a represents the detection before the split from the original bird. Ideally, 
  3_2b is the detection history that is then split upon to create 3_4a and 3_4b. However
  the split after time 4 would have created a vector of all 0's for individual 
  3_4b, which means it was technically not detected. So that split would've been
  a labelling error. That's why I ignore the effects of that split at time 4, and 
  instead of showing detection histories for 3_2a, 3_4a and 3_4b, I show the detection
  histories of 3_2a and 3_4 because individual 3_4b would not have been spotted.

```{r}
dfError = generateErrors(dfRegular,maxMinute,alpha,1)
#dfError$Detected = rowSums(dfError[,1:maxMinute])
head(dfError)
```

```{r}
dfError_Summarized =  dfError %>%
  group_by(locationID) %>%
  summarise(C_i_error = n())
dfError_Summarized = left_join(dfRegular_Summarized, dfError_Summarized, by= "locationID")
head(dfError_Summarized)
```


```{r}
head(dfError)
```


# Graphs

Reset parameter values. I changed the p_1m above slightly because I wanted to generate
enough test cases to see if the way I was creating errors worked. Now that it is, I'll
go back to using the regular p_1m in order to compare values now to analysis in
the previous .Rmd about regular point count data.


```{r}
p_1m = 0.1 # what was used in the previous .Rmd file
maxMinute = 5
psi=0.6
n_locations = 50
lambda = 15 # get this number from some data?
alpha = 0.3
```


## Generate Population Data
```{r}
set.seed(1)
locationID = 1:n_locations
#N_i = rpois(n_locations,lambda)
N_i = rnbinom(n_locations, mu = lambda, size= 1)
```


```{r}
# view results
result = returnData(N_i,p_1m,maxMinute, seed = 1)
dfRegular_Cumulative = result[[1]]
dfRegular = result[[2]]
dfRegular_Summarized = result[[3]]
dfRegular_Summarized$locationID = 1:n_locations


head(dfRegular_Cumulative)
head(dfRegular)
head(dfRegular_Summarized)
```


```{r}
resultError = returnErrors(N_i, p_1m, maxMinute, alpha,seed = 1)

dfError_cumulative = resultError[[1]] # cumulative counts
dfError = resultError[[2]] # incremental detections
dfError_Summarized = resultError[[3]] # summarized data

head(dfError_cumulative)
head(dfError)
head(dfError_Summarized)
```

## Compare Point counts

I will compare the distribution of errors to the population distribution and the 
distribution of correct counts. 

```{r}
dfError_Summarized%>%
  pivot_longer(cols= c("N_i","C_i","C_i_error"),names_to = "Type", values_to= "Values")%>%
  ggplot()+
  geom_density(aes(x=Values, color= Type))+
  theme_minimal()+ 
  labs(title= "Distribution of Population, Point Counts, and Erroneous Point Counts")
```


Facet for easier comparison. Looks like erroneous point count data is still somewhat
similar to the distribution of C_i, but the density at certain points is lower. The 
next step is to generate distributions for various values of alpha - this is a good 
way to check that error mechanism is working since alpha = 0 should give me the same
distribution as C_i.
```{r}
dfError_Summarized%>%
  pivot_longer(cols= c("N_i","C_i","C_i_error"),names_to = "Type", values_to= "Values")%>%
  ggplot()+
  geom_density(aes(x=Values, fill= Type))+
  facet_wrap(~Type)+
  theme_minimal()+ 
  labs(title= "Distribution of Population, Point Counts, and Erroneous Point Counts")
```


A list of alpha values are generated and the result of C_i error are shown for each 
value of alpha.
```{r}
lstAlpha = c(0,0.3,0.5,0.7,0.9)
dfAlpha = data.frame(N_i= N_i,
                     C_i= dfRegular_Summarized$C_i)

# loop to create new C_i_errors based on alpha
for(alpha in lstAlpha){
  resultError = returnErrors(N_i, p_1m, maxMinute, alpha,seed = 1)
  dfError_Summarized = resultError[[3]] # summarized data
  
  newColName = paste0("C_i_error_alpha_",alpha)
  dfAlpha[,newColName] = dfError_Summarized$C_i_error
}
```


```{r}
head(dfAlpha)
```

Need to zoom in to see image properly, but graphs right to me. Ci and C_i_error_alpha_0
have the same distribution (as it should be). We can see that as the probability of errors
increases, the distribution has a flatter peak and looks a bit more similar 
```{r}
dfAlpha%>%
  pivot_longer(cols= everything() ,names_to = "Type", values_to= "Values")%>%
  ggplot()+
  geom_density(aes(x=Values, fill= Type))+
  facet_wrap(~Type)+
  theme_minimal()+ 
  labs(title= "Distribution of Population, Point Counts, and Erroneous Point Counts")
```

Looking at it unfaceted is interesting because you can see how the peak becomes smaller
and the median seems to shift as well. Maybe a boxplot to illustrate would help.
```{r}
dfAlpha%>%
  pivot_longer(cols= everything() ,names_to = "Type", values_to= "Values")%>%
  ggplot()+
  geom_density(aes(x=Values, color= Type))+
  theme_minimal()+ 
  labs(title= "Distribution of Population, Point Counts, and Erroneous Point Counts")
```
I think boxplot highlights this well - we see the shifting median and increasing 
outliers as alpha rises in value.
```{r}
dfAlpha%>%
  pivot_longer(cols= everything() ,names_to = "Type", values_to= "Values")%>%
  filter(Type != "C_i_error_alpha_0")%>%
  ggplot()+
  geom_boxplot(aes(x=Type,y=Values, color= Type))+
  theme_minimal()+ 
  labs(title= "Distribution of Population, Point Counts, and Erroneous Point Counts")+
  theme(axis.text.x = element_blank())
```




